---
title: "Warren_Navpreet_turn_in"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Gathering data sets for NYC zipcodes and asc data (which consists of the SES features)
First, we have to insert our libraries, zip codes, and data files to make sure we have NYC zip codes and SES data.
```{r cars}
library(tidyverse)
library(tidycensus)
library(tigris)
#uncomment below and run it in case there's an error with tibbles
#install.packages("sf")
library(sf)
options(tigris_use_cache = TRUE)

load("census_2016_data.RData")

#in case you need this
#census_api_key("KEY_GOES_HERE", install = TRUE)

#grab the zip codes for NYC
# came from this link https://github.com/nychealth/coronavirus-data/blob/097cbd70aa00eb635b17b177bc4546b2fce21895/tests-by-zcta.csv
zippy <- read.csv("April_1_pos_tests.csv")
zippy <- zippy %>% mutate(GEOID = MODZCTA)
zippy <- zippy %>% mutate(prop_COVID = Positive/Total)
```

#Comparing summary statistics with the report
We started off by comparing the medians and confidence intervals for each of the variables, plotting the data on a map, and create a univariate regression model with the results from the paper.

We grab the proportion of 18-64 year old population that is uninsured. We'll make a table, a linear model and summarize all of that data to compare the results from the paper. 
```{r pressure, echo=FALSE}
#proportion of 18-64 year old population that is uninsured

#make zip code numeric, to match type with the other one
df_uninsured$GEOID <- as.numeric(df_uninsured$GEOID)

#join to keep the NYC zip codes
df_uninsured <-  inner_join(df_uninsured, zippy, by="GEOID")

#get rid of extra columns from zippy
df_uninsured <- df_uninsured[, -c(11:13)]

#create a column with proportion
df_uninsured <- df_uninsured %>% mutate(pop_uninsured = (pop_18_to_34E + pop_35_to_64E)/(tot_18_to_34E + tot_35_to_64E))

#median and confidence intervals
median(df_uninsured$pop_uninsured)
quantile(df_uninsured$pop_uninsured, 0.25)
quantile(df_uninsured$pop_uninsured, 0.75)
```
Results from  the median and quartile ranges for the uninsured population were consistent with the ones written in the paper. The median was 13.6% (8.6%-17.3%) in comparison to 13%(9-17%) from the paper.
```{r}
#graph it on the map of nyc
ggplot(data = df_uninsured, aes(fill = pop_uninsured)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1) +
  ggtitle("Proportion of 18-64 year olds that are uninsured")
```

```{r}
#make a linear model
uninsured_model <- lm(prop_COVID ~ pop_uninsured, data = df_uninsured)

#get summary
summary(uninsured_model)
```
The regression results showed a r-squared value of 34% in comparison to the 38% from the paper. 

Next we grab income. And do the same: make a map, a linear model, and a summary.
```{r}
#get income data

#make zip code numeric, to match type with the other one
med_inc$GEOID <- as.numeric(med_inc$GEOID)

#join to keep the NYC zip codes
med_inc <-  inner_join(med_inc, zippy, by="GEOID")

#get rid of extra columns
med_inc <- med_inc[, -c(5:7)]

#median and confidence intervals
median(med_inc$income)
quantile(med_inc$income, 0.25)
quantile(med_inc$income, 0.75)
```
The results for the median values were only a couple of thousands off from the paper (since our median did not take into account weight of each zip code like Jake mentioned). The median value was 60,526 (45,643-81,678).
```{r}
#graph on a map
#make columns to take in millions
med_inc <- med_inc %>% mutate(income_mil = income/1000000)

#graph
ggplot(data = med_inc, aes(fill = income_mil)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "YlGn", 
                       direction = 1) + 
  ggtitle("Median income (in millions 2016$)")
```

```{r}
#linear model, then summarise
income_model <- lm(prop_COVID ~ income_mil, data = med_inc)

summary(income_model)
```
The results for the univariate regression was 29% which was close to the results from the paper (32%). 

Now we do the same for proportion of people who identify as white.

```{r}
#rinse and repeat
race_white$GEOID <- as.numeric(race_white$GEOID)
race_white <-  inner_join(race_white, zippy, by="GEOID")
race_white <- race_white[, -c(7:9)]

#divide to get portion
race_white <- race_white %>% mutate(prop_white = whiteE / totE)

median(race_white$prop_white)
quantile(race_white$prop_white, 0.25)
quantile(race_white$prop_white, 0.75)
```
The results for this section were also similar to the ones presented in the paper with a median of 49.4% (22.6%-70%).

```{r}
ggplot(data = race_white, aes(fill = prop_white)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "Purples", 
                       direction = 1) +
  ggtitle("Proportion self-identify as White")
```

```{r}
#linear regression model
race_model <- lm(prop_COVID ~ prop_white, data = race_white)

summary(race_model)
```
For regression the r-squared value was 33.52% which is very close to the 34% mentioned in the paper. 

Now we do the same for house size. Note this one has more calls since the data is very segmented.

```{r}
#same as the previous
house_size$GEOID <- as.numeric(house_size$GEOID)
house_size <-  inner_join(house_size, zippy, by="GEOID")

house_size <- house_size[, -c(21:23)]

#get proportions
house_size <- house_size %>% mutate(prop_four_up =  (fourE + fiveE + sixE + sev_moreE +
                                    nfamfourE + nfamfiveE +nfamsixE + nfamsev_moreE) / totalE )

median(house_size$prop_four_up)
quantile(house_size$prop_four_up, 0.25)
quantile(house_size$prop_four_up, 0.75)
```
Results for the proportion in households of 4 or more showed a median of 24%(15%-31%), in comparison to the 
```{r}
ggplot(data = house_size, aes(fill = prop_four_up)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1) +
  ggtitle("Proportion in households of 4 or more")
```

```{r}
house_model <- lm(prop_COVID ~ prop_four_up, data = house_size)

summary(house_model)
```
 The r-squared value from the regression model was 

Now the same for people taking the bus.

```{r}
pub_trans$GEOID <- as.numeric(pub_trans$GEOID)
pub_trans <-  inner_join(pub_trans, zippy, by="GEOID")

pub_trans <- pub_trans[, -c(7:9)]

pub_trans <- pub_trans %>% mutate(prop_bus = busE/totalE)

median(pub_trans$prop_bus)
quantile(pub_trans$prop_bus, 0.25)
quantile(pub_trans$prop_bus, 0.75)
```

```{r}
ggplot(data = pub_trans, aes(fill = prop_bus)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1) + 
  ggtitle("Proportion of population that commutes by bus")
```

```{r}
bus_model <- lm(prop_COVID ~ prop_bus, data = pub_trans)
summary(bus_model)
```


Now for the elderly population

```{r}
#this one we have to combine a lot, since they do it in male and female, in increments of a couple years
elderly$GEOID <- as.numeric(elderly$GEOID)
elderly <-  inner_join(elderly, zippy, by="GEOID")
elderly <- elderly[, -c(29:31)]

elderly <- elderly %>% mutate(prop_eld =  (m65_66E + m66_67E + m67_68E + m68_69E + m69_70E+ m70aE +
                                           f65_66E + f66_67E + f67_68E + f68_69E + f69_70E+ f70aE)/totalE )

median(elderly$prop_eld)
quantile(elderly$prop_eld, 0.25)
quantile(elderly$prop_eld, 0.75)
```

```{r}
ggplot(data = elderly, aes(fill = prop_eld)) + 
  geom_sf() + 
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1) +
  ggtitle("Proportion of population 65+ years of age")
```

```{r}
elderly_model <- lm(prop_COVID ~ prop_eld, data = elderly)
summary(elderly_model)
```

Now for Table 1. Where we make a linear model of 4 of the previous variables.

```{r}
#unisured, white, over3, income

#we must drop the geometry columns b/c you can't join with those columns
#comment in/out if needed
df_uninsured <- st_drop_geometry(df_uninsured)
race_white <- st_drop_geometry(race_white)

merged_table <- inner_join(df_uninsured, race_white, by="GEOID")

#comment in/out if needed
house_size <- st_drop_geometry(house_size)
med_inc <- st_drop_geometry(med_inc)
med_inc$GEOID <- as.numeric(med_inc$GEOID)
second_merged <- inner_join(house_size, med_inc, by="GEOID")

#merge everything for the linear model
merged_table$GEOID <- as.numeric(merged_table$GEOID)
full_merge <- inner_join(merged_table, second_merged, by="GEOID")

full_lm <- lm(prop_COVID.x.x ~ prop_four_up + pop_uninsured + prop_white + income_mil, data = full_merge)

#here are the confidence intervals and estimates
confidence_table1 <- confint(full_lm, level = 0.95)
estimates_table1 <- summary(full_lm)$coefficients

#Cleaner Table to see results
final_table1 <- data.frame(col1 = c('Intercept', 'Proportion of population living in households with more than 3 inhabitants', 'Proportion of the 18- to 64- year old population that is uninsured', 'Proportion of the population that self-identified their race as white', 'Median Household Income (2016$)'),
                           col2 = c(0.46, 0.33, 0.25, -0.09, -0.34),
                           col3 = c(0.41, 0.23, 0.04, -0.14, -0.76),
                           col4 = c(0.52, 0.44, 0.47, -0.04, 0.07)
)
colnames(final_table1) <- c('Variables', 'Estimate', '2.5%', '97.5%')
final_table1
```

R-Squared value
```{r}
summary(full_lm)$r.squared
```

#Second Plots 
Now for plot 2. Where we grab safegraph data and find the trend in change of mobility.
```{r}
load('/data/safegraph/safegraph.Rdata')

zippy <- read.csv("April_1_pos_tests.csv")
zippy <- zippy %>% mutate(postal_code = MODZCTA)

#zippy$postal_code <- as.numeric(zippy$postal_code)
safegraph <- inner_join(safegraph, zippy, by="postal_code")

safegraph <- safegraph[, -c(7:9)]

#to get the baseline, grab february and grab median of average visits per day
baseline_df <- safegraph %>% filter(grepl("^2020-02-", date))

baseline_df <- baseline_df %>% group_by(postal_code) %>% summarize(base_med = median(avg_visits_per_day))

#join the median of that zip code to the main table
safegraph <- inner_join(safegraph, baseline_df, by= "postal_code") 

#get the difference from median
safegraph<- safegraph %>% mutate(change_base = (avg_visits_per_day - base_med) / base_med)

#make date into a factor for the graph
safegraph$date <- as.factor(safegraph$date)

#get rid of feb since it wasn't in the paper
safegraph <- safegraph %>% filter(grepl("^2020-03-", date) | grepl("^2020-04-", date))

#get median of all zips
dates_listy <- safegraph %>% group_by(date) %>% summarise(all_zip_med = median(change_base, na.rm=TRUE), 
                                                          all_zip_quant_low = quantile(change_base, .25 ,na.rm=TRUE),
                                                          all_zip_quant_high = quantile(change_base, .75 ,na.rm=TRUE))

#plot
ggplot() +
  geom_violin(data = safegraph, aes(x = change_base, y = date), color = "orange") +
  geom_pointrange(data = dates_listy, aes(x = all_zip_med, y = date, xmin = all_zip_quant_low, xmax = all_zip_quant_high), color = "red") +
  xlim(-1,2) + 
  xlab("Change in mobility relative to baseline") + 
  ylab("Date")

```

#Table 2 Replications
And finally for table 2. Where we include everything and also grab the mobility data
```{r}
#note: we already got rid of the geometry columns for several of the tables
#i left them in comments in case you want to test this without running the table 1 block

#df_uninsured <- st_drop_geometry(df_uninsured)
#race_white <- st_drop_geometry(race_white)

mergerd_table <- inner_join(df_uninsured, race_white, by="GEOID")

#house_size <- st_drop_geometry(house_size)
#med_inc <- st_drop_geometry(med_inc)

second_merged <- inner_join(house_size, med_inc, by="GEOID")
full_merge <- inner_join(mergerd_table, second_merged, by="GEOID")
elderly <- st_drop_geometry(elderly)
pub_trans <- st_drop_geometry(pub_trans)
third_merged <- inner_join(elderly, pub_trans, by = "GEOID")
six_merge <- inner_join(third_merged, full_merge, by="GEOID")
full_lm2 <- lm(prop_COVID.x.y ~ prop_eld + prop_bus + income_mil + prop_white + pop_uninsured + prop_four_up, data = six_merge)


#here are the confidence intervals and estimates -
no_mob <- confint(full_lm2)
no_mob_summ <- summary(full_lm2)

no_mob
no_mob_summ

#Cleaner Table to see results
final_table2a <- data.frame(col1 = c('Intercept', 'Proportion of population that is 65 or older', 'Proportion of population using public transportation to commute to work that includes bus travel', 'Median Household Income (2016$)', 'Proportion of the population that self-identified their race as white', 'Proportion of the 18- to 64- year old population that is uninsured', 'Proportion of population living in households with more than 3 inhabitants'),
                           col2 = c(0.44, 0.21, -0.11, -0.28, -0.11, 0.27, 0.37),
                           col3 = c(0.37, -0.01, -0.31, -0.72, -0.16, 0.04, 0.24),
                           col4 = c(0.52, 0.44, 0.09, 0.17, -0.06, 0.49, 0.49)
)
colnames(final_table2a) <- c('Variables', 'Estimate', '2.5%', '97.5%')
final_table2a
```

R-Squared Value
```{r}
summary(full_lm2)$r.squared
```

```{r}
#now with mobility
#first change postal code to GEOID
safegraph <- safegraph %>% rename(GEOID = postal_code)
safegraph$GEOID <- as.numeric(safegraph$GEOID)
safegraph <- safegraph %>% filter(grepl("2020-04-01", date))

seven_merge <- inner_join(six_merge, safegraph, by = "GEOID")

#drop nas and infs
seven_merge <- na.omit(seven_merge)
seven_merge <- seven_merge %>% filter(change_base != Inf)

seven_lm <- lm(prop_COVID.x.y ~ prop_eld + prop_bus + income_mil + prop_white + pop_uninsured + prop_four_up + change_base, data = seven_merge)

#table 2
mobility <- confint(seven_lm)
mobility_summ <- summary(seven_lm)

mobility
mobility_summ

#Cleaner Table to see results
final_table2b <- data.frame(col1 = c('Intercept', 'Proportion of population that is 65 or older', 'Proportion of population using public transportation to commute to work that includes bus travel', 'Median Household Income (2016$)', 'Proportion of the population that self-identified their race as white', 'Proportion of the 18- to 64- year old population that is uninsured', 'Proportion of population living in households with more than 3 inhabitants','Mobility'),
                           col2 = c(0.45, 0.19, -0.11, -0.52, -0.09, 0.26, 0.36, -0.01),
                           col3 = c(0.37, -0.04, -0.32, -1.02, -0.14, 0.03, 0.23, -0.05),
                           col4 = c(0.53, 0.42, 0.09, -0.02, -0.04, 0.48, 0.49, 0.02)
)
colnames(final_table2b) <- c('Variables', 'Estimate', '2.5%', '97.5%')
final_table2b
```

R-squared value
```{r}
summary(seven_lm)$r.squared
```

#Extension
We extended the replication project to cover the entire New York State instead of New York City to see if there was a difference in the prediction of positive covid cases. The method utilized was the same as the replication just for a larger data set. For this case, we utilized only the 6 SES variables for prediction. 
```{r}
load("census_2016_data.RData")
#uninsured
df_uninsured$GEOID <- as.factor(df_uninsured$GEOID)
df_uninsured <- df_uninsured %>% mutate(pop_uninsured = (pop_18_to_34E + pop_35_to_64E)/(tot_18_to_34E + tot_35_to_64E))

#income data
med_inc$GEOID <- as.factor(med_inc$GEOID)
med_inc <- med_inc %>% mutate(income_mil = income/1000000)

#race white
race_white$GEOID <- as.factor(race_white$GEOID)
race_white <- race_white %>% mutate(prop_white = whiteE / totE)

#house size
house_size$GEOID <- as.factor(house_size$GEOID)
house_size <- house_size %>% mutate(prop_four_up =  (fourE + fiveE + sixE + sev_moreE +
                                                       nfamfourE + nfamfiveE +nfamsixE + nfamsev_moreE) / totalE )

#busing
pub_trans$GEOID <- as.factor(pub_trans$GEOID)
pub_trans <- pub_trans %>% mutate(prop_bus = busE/totalE)

#elderly
elderly$GEOID <- as.factor(elderly$GEOID)
elderly <- elderly %>% mutate(prop_eld =  (m65_66E + m66_67E + m67_68E + m68_69E + m69_70E+ m70aE +
                                             f65_66E + f66_67E + f67_68E + f68_69E + f69_70E+ f70aE)/totalE )

#drop the geometry columns b/c you can't join, then join everything
df_uninsured <- st_drop_geometry(df_uninsured)
race_white <- st_drop_geometry(race_white)

mergerd_table <- inner_join(df_uninsured, race_white, by="GEOID")

house_size <- st_drop_geometry(house_size)
med_inc <- st_drop_geometry(med_inc)

second_merged <- inner_join(house_size, med_inc, by="GEOID")
full_merge <- inner_join(mergerd_table, second_merged, by="GEOID")

elderly <- st_drop_geometry(elderly)
pub_trans <- st_drop_geometry(pub_trans)

third_merged <- inner_join(elderly, pub_trans, by = "GEOID")
six_merge <- inner_join(third_merged, full_merge, by="GEOID")

zippy2 <- read.csv("apr_1st_NY.csv")
zippy2$Zip_Code <- as.factor(zippy2$Zip_Code)
zippy2 <- zippy2 %>% mutate(GEOID = Zip_Code)
zippy2 <- zippy2 %>% mutate(prop_COVID = Positive_Cases/Total_Tests)

seven_merge <- inner_join(six_merge, zippy2, by = "GEOID")

full_lm <- lm(prop_COVID ~ prop_eld + prop_bus + income_mil + prop_white + pop_uninsured + prop_four_up, data = seven_merge)

#here are the confidence intervals and estimates -
confint(full_lm)
summary(full_lm)
```
Conclusion & Further Work
We are not certain as to why the coefficients are consistenly larger when including most zipcodes from New York State in comparison to New York City. Our best guess is that because all of NYC, the living spaces are close quarters that it may bias the model because it is unaccounted for. Whereas, when we included zipcodes from all of New York State, that includes many more town structures that aren't just close living spaces and not necessarily heavily used transportation areas. Therefore, it biases the data less since those types of towns are being accounted for and the factors that we measured are being able to show a more accurate representation of their influence. In summary we think it's because NYC has closer housing and heavily used tranportation which is not true for all of New York State. 

We would like to further explore this question after adding mobility data as well as thinking of methods in which we can capture information like zipcodes that have much more transportation usage and including the weight of those factors into the model.










